Digitization of the Australian Parliamentary Debates, 1998-2022

### 2023-04-05 paper/paper-nature_format.qmd
updated zenodo link
Update the paper

---
title: "Digitization of the Australian Parliamentary Debates, 1998-2022"
number-sections: true
format: 
  pdf:
    keep-tex: true
bibliography: references.bib
editor: visual
date: "now"
date-format: long
author:
  - Lindsay Katz^[University of Toronto, lindsay.katz@mail.utoronto.ca]
  - Rohan Alexander^[University of Toronto, rohan.alexander@utoronto.ca]
abstract: "Public knowledge of what is said in parliament is a tenet of democracy, and a critical resource for political science research. In Australia, following the British tradition, the written record of what is said in parliament is known as Hansard. While the Australian Hansard has always been publicly available, it has been difficult to use for the purpose of large-scale macro and micro-level text analysis because it has only been available as PDFs or XMLs. Following the lead of the Linked Parliamentary Data project which achieved this for Canada, we provide a new, comprehensive, high-quality, rectangular database that captures all proceedings of the Australian parliamentary debates from 1998 to the present using Hansard. The database is publicly available and can be linked to other datasets such as election results. The creation and accessibility of this database enables the exploration of questions that have not been possible to explore, serving as a valuable resource for both researchers and policymakers."
---

@@

The official written record of parliamentary debates, formally known as Hansard, plays a fundamental role in capturing the history of political proceedings and facilitating the exploration of valuable research questions. Originating in the British parliament, the production of Hansard became tradition in many Commonwealth countries such as Canada, the United Kingdom, and Australia [@vice2017history]. Given the content and magnitude of these records, they have significance, particularly in the context of political science research. In the case of Canada, the Hansard has been digitized for 1901 to 2019 [@lipad]. Having a digitized version of Hansard enables researchers to perform advanced analyses on these records using text analysis tools and statistical modelling. Following the lead of that project, in this paper we introduce a novel database for the Australian Hansard. This is composed of individual datasets for each sitting day in the House of Representatives from 1998 to 2022, containing details on everything said in parliament in a form that can be readily used by researchers. With the development of tools for large-scale text analysis, this database will serve as a valuable resource for efficiently studying and exploring political behaviour in Australia over time.

The Australian House of Representatives performs a number of crucial governmental functions, such as creating new laws and overseeing government expenditure [@australia2018house, ch. 1]. The House operates under a parallel chamber setup, meaning there are two debate venues where proceedings take place: the Chamber, and the Federation Chamber. Sittings of the House follow a predefined order of business, regulated by procedural rules called standing orders [@australia2018house, ch. 8]. A typical sitting day in the Chamber has a number of scheduled proceedings including debates on government business, 90 second member statements, and Question Time [@australia2018house, ch. 8]. The Federation Chamber was created in 1994 as a subordinate debate venue of the Chamber. This allows for better time management of House business as its proceedings occur simultaneously with those of the Chamber [@australia2018house, ch. 21]. Sittings in the Federation Chamber are quite different to those of the Chamber in terms of their order of business and scope of discussion. Business matters discussed in the Federation Chamber are limited largely to intermediate stages of bill development, and the business of private Members [@australia2018house, ch. 21]. It is the recording and compilation of all of these proceedings on which Hansard is based, and it is essentially, but not entirely, verbatim.

A week or so after each sitting day, a transcript is available for download from the official Parliament of Australia website in both PDF and extensible markup language (XML) form. The PDF is the official release. For early decades there is a need to digitize PDFs using Optical Character Recognition (OCR) technology and there are many flaws [@sherratt]. The PDF imposes formatting designed for humans to read with ease, whereas XML is designed for consistency and machine legibility. The nature of XML enables us to more easily use code to manipulate these records at scale, motivating our choice to develop our database solely using the XML formatted files. In cases where we were unsure on how to proceed with processing the XML, we defer first to the PDF, and then to the video recording of the proceeding, if available.

@@

Many papers exist which use components of Australian Hansard to explore various research topics. For example, @salisbury2011 used online Hansard records to investigate the occurrences of unparliamentary comments by members of the House, where the Speaker tells that member to withdraw their remark. @rasiah2010framework worked with Question Time data from Hansard transcripts during February and March of 2003, to investigate resistance of politicians in answering questions about Iraq. In their work, @fraussen2018 use Hansard to quantify political prominence by investigating strategic mentions of interest groups by elected officials. Finally, the work of @alexander2021 offers a dataset of the Australian Hansard, along with an analysis of the effect of elections and changes in Prime Ministers upon topics mentioned in parliament. @alexander2021 created this database with the static PDF versions of Hansard, using OCR to digitize these files into text which is suitable for analysis. This means there are considerable digitization errors especially in the first half of the dataset.

While there is evidently a growing body of literature on this topic, there is still no comprehensive database for Australian Hansard based on XML that spans from 1901 to the present day. Our work serves to bridge this gap.

@@

Our database contains one comma-separated value (CSV) file for each sitting day of the House of Representatives from 02 March 1998 to 08 September 2022. We developed four scripts to produce all of these files, where each script is designed to parse Hansard documents from a specific portion of the 1998 to 2022 time frame.

This section will be structured as follows. First, we will provide an overview of our approach to understanding and parsing an individual Hansard XML document, which informed all four scripts used to create our database. This will be supplemented with an excerpt from a Hansard XML to provide a visual example of its structure. Next we will explain the specific differences between the four scripts, and outline what structural changes necessitated their separate development. We will then provide details on the methodological intricacies of three core components of Hansard proceedings: Question Time, interjections, and stage directions. Finally, we will discuss the script we developed to fill in remaining missing speaker details, which each file in our database was passed to after being parsed and cleaned.

@@

The approach to parsing contents of an XML document is heavily dependent on its tree structure. As such, to create this database, we started by looking at a single Hansard XML transcript from 2019. Doing so enabled us to identify the various components of interest in the document, and how each one can be parsed according to its corresponding structural form. Parsing was performed in `R` using the `XML` and `xml2` packages [@XML_pkg; @xml2_pkg]. Focusing on one transcript also allowed us to ensure that all key components of the transcript were being parsed and captured with as much detail as possible. The typical form of a Hansard XML transcript is summarized in the nested list below. Note that this provides a broad overview of the structure, and does not contain every possible nested element that may be found in a Hansard XML.

@@

The outer-most node, also known as the parent node, is denoted `<hansard>` and serves as a container for the entire document. This parent node may have up to four child-nodes, where the first child-node contains details on the specific sitting day. Next, `<chamber.xscript>` contains all proceedings of the Chamber, `<fedchamb.xscript>` contains all proceedings of the Federation Chamber, and `<answers.to.questions>` contains Question Time proceedings. Note that the Federation Chamber does not meet on every sitting day, so this child element is not present in every XML file. The use of separate child nodes allows for the distinction of proceedings of the Chamber and Federation Chamber. The structure of the `<chamber.xscript>` and `<fedchamb.xscript>` nodes are generally the same, where the proceeding begins with a business start which is followed by a series of debates. Debate nodes can contain a `<subdebate.1>` child-node which has a `<subdebate.2>` child-node nested within it. That said, sometimes `<subdebate.2>` is not nested within `<subdebate.1>`. Each of these three elements (i.e. `<debate>`, `<subdebate.1>`, and `<subdebate.2>`) as well as their respective sub-elements contain important information on the topic of discussion, who is speaking, and what is being said. The `<speech>` node within each one contains the bulk of the text associated with that debate or sub-debate. A typical `<speech>` node begins with a `<talk.start>` sub-node, providing information on the speaker and the time of their first statement. Unsurprisingly, speeches rarely go uninterrupted in parliamentary debate settings --- they are often composed of a series of interjections and continuations. These statements are categorized under different sub-nodes depending on their nature, such as `<interjection>` or `<continuation>`. The final key component of Hansard is Question Time, in which questions and answers are classified as unique elements. More detail on the purpose and processing of Question Time will follow in @sec-qa.

@fig-xml1 provides an example of the beginning of an XML file for Hansard, which illustrates the structure outlined in the nested list above. As stated, the XML structure begins with a parent element `<hansard>` (highlighted in blue), followed by a child element `<session.header>` (highlighted in yellow) with sub-child elements such as the date and parliament number, which are all highlighted in pink. Next there is the child element containing everything that takes place in the Chamber, `<chamber.xscript>`, which is also highlighted in yellow in @fig-xml1. As previously mentioned, the first sub-node of `<chamber.xscript>` is the business start. The structure of the business start can be seen between the nodes highlighted in green in @fig-xml1, where the content we parse from the business start is highlighted in orange.

@@

Evidently, the nature of XML formatting means that different pieces of information are categorized under a series of uniquely named and nested nodes. As a result, to parse each piece of information, one must specify the unique hierarchy of nodes in which it is structured. This is known as an XPath expression, and tells the parser how to navigate the XML document to obtain the desired information. For example, the session header date in @fig-xml1 can be accessed using the XPath expression "`hansard/session.header/date`". When specifying an XPath expression, one can use an "or" operator to obtain elements from multiple node paths at once, in the order that they appear in the document. We did so throughout our script as we parsed uniquely nested speech content, as it allows for the correct ordering of elements to be maintained. We began our first script by parsing all the business start, speech text, and Question Time contents contained in the XML document, using these unique XPath expressions to do so. Hansard records contain a number of statements throughout that were not explicitly spoken in parliament. For example, the text in @fig-content_ex beneath the bold subtitle "Report from Federation Chamber" is not attributed to a particular member and was not spoken aloud during the parliamentary proceedings. In contrast, the text which comes afterward was spoken aloud in parliament by The Speaker at 14:01. Generally, unspoken text is nested under `<debate.text>` or `<subdebate.text>` nodes, while spoken text is categorized under `<speech>`. Importantly, in our work we focus primarily on parsing spoken speech content (i.e. what is nested within `<speech>` nodes), such as what is stated by The Speaker in @fig-content_ex. For this reason, many Hansard transcriptions which are not spoken text may not be captured in our database. In future work, capturing this unspoken content in our database would be a priority.

The next step was to further develop our script to produce tidy data sets containing all parsed text elements, where each statement is separated onto its own row with details about the speaker, and rows are maintained in chronological order. This first involved correcting the variable classes and adding several indicator variables to differentiate where statements came from, such as Chamber versus Federation Chamber or sub-debate 1 versus sub-debate 2. The next key task stemmed from the fact that the raw text data were not separated by each statement when parsed. In other words, any interjections, comments made by The Speaker or Deputy Speaker and continuations within an individual speech were all parsed together as a single string. As such, the name, name ID, electorate and party details were only provided for the person whose turn it was to speak. Splitting up these speeches in a way which would be generalizable across sitting days required much thought and effort. @sec-interject will provide further details on the intricacies of this task.

Since we are looking at a wide time span of documents, there are many changes in the way they are formatted over time. These became apparent as we ran our script on XML files from earlier sitting days. Some changes are as subtle as a differently named child-node, while others are as extensive as a completely different nesting structure. Smaller changes were accounted for as we became aware of them, and embedded into the code in a way that would not cause issues for parsing more current Hansards with subtle differences in formatting. However, as mentioned, more significant changes in the XML structure of Hansard are what necessitated the development of separate scripts as we worked backwards in time. Further, while working backward in time, it became clear that not every sitting day contains every possible XML element. For example, some days did not have sub-debate 2 content, and some days did not have a Federation Chamber proceeding. To improve the generalizability of these scripts, if-else statements were embedded within the code wherever an error might arise due to a missing element. For example, the entire Federation Chamber block of code is wrapped in an if-else statement for each script, so that it only executes if what the code attempts to parse exists in the file.

Once the script ran without error for a few recent years of Hansard, we continued to work backwards until extensive changes in tree structure made our script incompatible with parsing earlier XML files. The earliest sitting day our first script can successfully parse is 14 August 2012. Before developing new scripts to parse earlier Hansard documents, we decided to prioritize cleaning and finalizing what we have been able to parse. As such we continued building our script, fixing any problems we noticed in the resulting datasets such as excess whitespace or spacing issues, and splitting up any additional sections of the parsed text onto separate rows where necessary. Specifically, we added a section of our script to separate out general stage directions. More information on this separation will be provided in @sec-stage. After completing our first script, it was formatted as a function which takes a single file name argument and produces one CSV file containing data on all proceedings from the given sitting day.

@@

As previously mentioned, we developed a total of four scripts to parse the 1998-2022 time frame of Hansard documents. Two main factors motivated us to create four scripts as opposed to just one, the first being structural variation in XML over time, and the second being improved computational efficiency with separate scripts. While all four scripts use the same general approach to parsing described in @sec-overview and produce the same CSV structure, the first and second scripts use a different method of data processing than the third and fourth scripts.

@@

While the modifications needed to develop the second script were quite straightforward, this was not the case for our next script. The typical tree structure of Hansard XMLs spanning from 1998 to March 2011 has an important difference from that of XMLs released after March 2011, necessitating many changes to be made in our methodology. In XMLs after March 2011, which our first two scripts successfully parse, the first two child-nodes of `<speech>` are typically `<talk.start>`, and `<talk.text>`. The first child-node contains data on the person whose turn it is to speak, and the second contains the entire contents of that speech --- including all interjections, comments, and continuations. After the `<talk.text>` element closes, there are typically a series of other child-nodes which provide a skeleton structure for how the speech proceedings went in chronological order. For example, if the speech began, was interrupted by one member, and then continued uninterrupted until the end, there would be one `<interjection>` node and one `<continuation>` node following the `<talk.text>` node. These would contain details on the person who made each statement such as their party and electorate.

@@

This was not isolated to just the beginning of speeches --- details on individuals interjecting or commenting during speeches were also captured this way. To separate statements out correctly, we collected all of these patterns using the `<talk.start>` node, and used them to split statements wherever one of these patterns was found. After separating the statements, we were able to remove these patterns from the body of text. We also used this method of extracting and later removing unwanted patterns for other pieces of data which did not belong to the debate proceedings, such as sub-debate titles.

@@

Our approach to parse the `"chamber.xscript"` speeches used in all four scripts meant that all questions without notice content was already parsed in order. For the first two scripts, questions and answers were already separated onto their own rows. For the third and fourth scripts, just as we did with the rest of the speech content, we used those patterns of data preceding the text to separate questions and answers. Finally, since questions in writing exist in their own child node we were able to use the same parsing method for all scripts, which was simply to extract all question and answer elements from the `"answers.to.questions"` child node.

We then added binary flags to differentiate between questions and answers. To do this in the first and second scripts, we separately re-parsed question and answer content using the XPath expressions `"chamber.xscript//question"` and `"chamber.xscript//answer"`, added the correct question and answer flags accordingly, and then added those flags back to the main dataframe based on exact text matches. For the third and fourth scripts, we made use of the fact that the patterns preceding text transcribed under a question node were stored separately from those transcribed under an answer node. As a result, we could readily use those patterns to flag questions and answers correctly based on which list of patterns it belonged to. Sometimes, questions were incorrectly transcribed under an answer node and vice-versa, in which cases we manually corrected the question and answer flags using the `str_detect` function from the `stringr` package. For instance, we check for any statements flagged as questions which include the phrase "has provided the following answer to the honourable member's question", in which case we re-code that statement as an answer.

The next step was to merge Question Time contents with all the debate speech. Doing so with the questions without notice content was straightforward. As mentioned, our method of parsing meant that everything was already parsed together in order, so we did not have to perform any additional merging. For questions in writing, merging this content was also straightforward due to the fact that it is always at the very end of Hansard. This means that we could simply bind question in writing rows to the bottom of the main dataframe. This approach was used for all four scripts.

@@

Below is an example of part of a speech we would need to split, extracted from Hansard on 30 November 2021, where Mr. van Manen is interrupted by The Speaker who states that the time for members' statements has concluded.

@@

As mentioned in @sec-diff, more recent years of Hansard XMLs contain a series of child-nodes which exist to capture the structure of interruptions in that speech. @fig-xml3 provides an example of this, where the speech was interrupted by a comment from The Deputy Speaker, and then the member continued their speech. Looking at the element names highlighted in blue, these child-nodes do not contain the actual text for the interjection or continuation --- this text is embedded within the speech above it. However, as shown by the content highlighted in pink in @fig-xml3, we were able to extract useful details on the individual interjecting which we could use later. Making use of this structure, we extracted names and information of all individuals that were categorized within the XML as interjections. We stored this as a dataframe called `"interject"`.[^1]

[^1]: We decided not to include this data in our final database, as it is all embedded in our resulting datasets which have a flag for interjections.

@@

We then created lists using both the `interject` and `main` dataframes to capture all the names of individuals who spoke that day. We added the names of all Members in a number of unique formats, due to the frequent variation in how names are transcribed in Hansard. When a Member interjects or continues a speech, the usual form of their name is a title followed by their first name or first initial and/or last name. There is also variation in the capitalization of these names.[^2] Another source of variation is in individuals with more than one first name, as sometimes only their initial first name is written, while other times their entire first name is written. Additionally, some surnames have punctuation, and some surnames have specific capitalization such as "McCormack", where even in full capitalization, the first "c" remains lower case. This variation demands careful consideration when writing regular expression patterns to search for in the text. In these lists we also accounted for any general interjection statements transcribed that were not attributed to an individual, such as "An opposition member interjecting-".

@@

Having these lists enabled us to extract the names of Members and their associated titles as they exist in the text, by searching for exact matches with regular expression patterns. We then used these extracted names to split all the speeches up, using regular expressions with lookarounds. A lookaround can be added to a regular expression pattern to enhance the specificity of matches. These were used to ensure that the text was not being split in the wrong places, such as places where Members were simply being named in the statement of another Member.

@@

Next, we merged our main dataframe with the lookup table to replace any incomplete names with their full names, and to fill in any gaps with available name ID, electorate, party, gender, and unique ID information. Finally, we were able to add a flag for interjections. Grouping our data by the speech number, we defined an interjection as a statement made by anyone who is not The Speaker, The Deputy Speaker, or the Member whose turn it was to speak. @fig-interject provides an example of a Federation Chamber proceeding with interjections. Evidently, statements made by the original speaker Stuart Robert or by The Deputy Speaker Maria Vamvakinou are not flagged as interjections.

@@

Having developed a successful methodology for splitting interjections, we used this to inform our general approach in the third and fourth scripts. However, the difference in data cleaning used in these scripts necessitated some departure from the original methodology. As discussed in @sec-diff, we used string patterns extracted from `<talk.start>` nodes to separate speeches. As evident in @fig-xml3, `<talk.start>` nodes are nested within `<interjection>` nodes, meaning that the patterns of data from interjection statements were separated out in the process. In other words, our approach to data cleaning in the third and fourth scripts separated out interjections in the process. This meant that we did not need to create lists of names and titles to search for in the text as we did before. However, we used the same list of general interjection statements to separate on that was used in the first two scripts. We then did an additional check for statements that may have not been separated due to how they were embedded in the XML, and separated those out where needed.[^3]

@@

>>>>

---
title: "Digitization of the Australian Parliamentary Debates, 1998-2022"
number-sections: true
format: 
  pdf:
    keep-tex: true
bibliography: references.bib
editor: visual
date: "now"
date-format: long
author:
  - Lindsay Katz^[University of Toronto, lindsay.katz@mail.utoronto.ca]
  - Rohan Alexander^[University of Toronto, rohan.alexander@utoronto.ca]
abstract: "Public knowledge of what is said in parliament is a tenet of democracy, and a critical resource for political science research. In Australia, following the British tradition, the written record of what is said in parliament is known as Hansard. While the Australian Hansard has always been publicly available, it has been difficult to use for the purpose of large-scale macro- and micro-level text analysis because it has only been available as PDFs or XMLs. Following the lead of the Linked Parliamentary Data project which achieved this for Canada, we provide a new, comprehensive, high-quality, rectangular database that captures proceedings of the Australian parliamentary debates from 1998 to 2022. The database is publicly available and can be linked to other datasets such as election results. The creation and accessibility of this database enables the exploration of new questions and serves as a valuable resource for both researchers and policymakers."
---

@@

The official written record of parliamentary debates, formally known as Hansard, plays a fundamental role in capturing the history of political proceedings and facilitating the exploration of valuable research questions. Originating in the British parliament, the production of Hansard became tradition in many Commonwealth countries such as Canada, the United Kingdom, and Australia [@vice2017history]. Given the content and magnitude of these records, they have significance, particularly in the context of political science research. In the case of Canada, the Hansard has been digitized for 1901 to 2019 [@lipad]. Having a digitized version of Hansard enables researchers to conduct text analysis and statistical modelling. Following the lead of that project, in this paper we introduce a similar database for Australia. This is composed of individual datasets for each sitting day in the House of Representatives from 1998 to 2022, containing details on everything said in parliament in a form that can be readily used by researchers. With the development of tools for large-scale text analysis, this database will serve as a resource for understanding political behaviour in Australia over time.

The Australian House of Representatives performs a number of crucial governmental functions, such as creating new laws and overseeing government expenditure [@australia2018house, ch. 1]. The House operates under a parallel chamber setup, meaning there are two debate venues where proceedings take place: the Chamber, and the Federation Chamber. Sittings of the House follow a predefined order of business, regulated by procedural rules called standing orders [@australia2018house, ch. 8]. A typical sitting day in the Chamber has a number of scheduled proceedings including debates on government business, 90 second member statements, and Question Time [@australia2018house, ch. 8]. The Federation Chamber was created in 1994 as a subordinate debate venue of the Chamber. This allows for better time management of House business as its proceedings occur simultaneously with those of the Chamber [@australia2018house, ch. 21]. Sittings in the Federation Chamber are different to those of the Chamber in terms of their order of business and scope of discussion. Business matters discussed in the Federation Chamber are limited largely to intermediate stages of bill development, and the business of private Members [@australia2018house, ch. 21]. It is the recording and compilation of these proceedings on which Hansard is based, and it is essentially, but not entirely, verbatim.

A week or so after each sitting day, a transcript is available for download from the official Parliament of Australia website in both PDF and extensible markup language (XML) form. The PDF is the official release. The PDF imposes formatting designed for humans to read with ease, whereas XML is designed for consistency and machine legibility. The nature of XML enables us to more easily use code to manipulate these records at scale, motivating our choice to develop our database solely using the XML formatted files. In cases where we were unsure on how to proceed with processing the XML, we defer first to the PDF, and then to the video recording of the proceeding, if available.

@@

Many papers exist which use components of Australian Hansard to explore various research topics. For example, @salisbury2011 used online Hansard records to investigate the occurrences of unparliamentary comments by members of the House, where the Speaker tells that member to withdraw their remark. @rasiah2010framework worked with Question Time data from Hansard transcripts during February and March of 2003, to investigate resistance of politicians in answering questions about Iraq. @fraussen2018 use Hansard to quantify political prominence by investigating strategic mentions of interest groups by elected officials. Finally, @alexander2021 construct a dataset of the Australian Hansard, along with an analysis of the effect of elections and changes in Prime Ministers upon topics mentioned in parliament. @alexander2021 created this database with the static PDF versions of Hansard, using OCR to digitize these files into text which is suitable for analysis. This means there are considerable digitization errors especially in the first half of the dataset.

While there is evidently a growing body of literature on this topic, there is still no comprehensive database for Australian Hansard based on XML that spans from 1901 to the present day. Our work begins to bridge this gap.

@@

Our database contains one comma-separated value (CSV) and one parquet file for each sitting day of the House of Representatives from 02 March 1998 to 08 September 2022. We developed four scripts to produce these files. Each script parses Hansard documents from a specific portion of the 1998 to 2022 time frame.

This section is structured as follows. First, we provide an overview of our approach to understanding and parsing an individual Hansard XML document, which informed the scripts used to create our database. This will be supplemented with an excerpt from a Hansard XML to provide a visual example of its structure. Next we will explain the specific differences between the scripts, and outline what structural changes necessitated their separate development. We then provide details on the methodological intricacies of three core components of Hansard proceedings: Question Time, interjections, and stage directions. Finally, we discuss the script we developed to fill in remaining missing speaker details, which each file in our database was passed to after being parsed and cleaned.

@@

The approach to parsing contents of an XML document depends on its tree structure. As such, to create this database, we started by looking at a single Hansard XML transcript from 2019. Doing so enabled us to identify the various components of interest in the document, and how each one can be parsed according to its corresponding structural form. Parsing was performed in `R` using the `XML` and `xml2` packages [@XML_pkg; @xml2_pkg]. Focusing on one transcript also allowed us to ensure that all key components of the transcript were parsed and captured in as much detail as possible. The typical form of a Hansard XML transcript is summarized in the nested list below. This provides an overview, but does not contain every possible nested element that may be found in a Hansard XML.

@@

The outer-most node, also known as the parent node, is denoted `<hansard>` and serves as a container for the entire document. This parent node may have up to four child-nodes, where the first child-node contains details on the specific sitting day. Next, `<chamber.xscript>` contains all proceedings of the Chamber, `<fedchamb.xscript>` contains all proceedings of the Federation Chamber, and `<answers.to.questions>` contains Question Time proceedings. The Federation Chamber does not meet on every sitting day, so this child element is not present in every XML file. The use of separate child nodes allows for the distinction of proceedings between the Chamber and Federation Chamber. The structure of the `<chamber.xscript>` and `<fedchamb.xscript>` nodes are generally the same, where the proceeding begins with `<business.start>` which is followed by a series of debates. Debate nodes can contain a `<subdebate.1>` child-node which has a `<subdebate.2>` child-node nested within it. That said, sometimes `<subdebate.2>` is not nested within `<subdebate.1>`. Each of these three elements (i.e. `<debate>`, `<subdebate.1>`, and `<subdebate.2>`) as well as their respective sub-elements contain important information on the topic of discussion, who is speaking, and what is being said. The `<speech>` node within each one contains the bulk of the text associated with that debate or sub-debate. A typical `<speech>` node begins with a `<talk.start>` sub-node, providing information on the speaker and the time of their first statement. Unsurprisingly, speeches rarely go uninterrupted in parliamentary debate settings --- they are often composed of a series of interjections and continuations. These statements are categorized under different sub-nodes depending on their nature, such as `<interjection>` or `<continuation>`. The final key component of Hansard is Question Time, in which questions and answers are classified as unique elements. More detail on the purpose and processing of Question Time will follow in @sec-qa.

@fig-xml1 provides an example of the beginning of an XML file for Hansard, which illustrates the structure outlined in the nested list above. As stated, the XML structure begins with a parent element `<hansard>` (highlighted in blue), followed by a child element `<session.header>` (highlighted in yellow) with sub-child elements such as the date and parliament number, which are all highlighted in pink. Next, there is the child element containing everything that takes place in the Chamber, `<chamber.xscript>`, which is also highlighted in yellow in @fig-xml1. As previously mentioned, the first sub-node of `<chamber.xscript>` is `<business.start>`. The structure of this can be seen between the nodes highlighted in green in @fig-xml1, where the content we parse from the business start is highlighted in orange.

@@

Evidently, the nature of XML formatting means that different pieces of information are categorized under a series of uniquely named and nested nodes. As a result, to parse each piece of information, one must specify the unique hierarchy of the nodes in which it is structured. This is known as an XPath expression, and tells the parser how to navigate the XML document to obtain the desired information. For example, the session header date in @fig-xml1 can be accessed using the XPath expression "`hansard/session.header/date`". When specifying an XPath expression, one can use an "or" operator to obtain elements from multiple node paths at once, in the order that they appear in the document. We did so throughout our script as we parsed uniquely nested speech content. This allows the correct ordering of elements to be maintained. We began our first script by parsing all the business start, speech text, and Question Time contents contained in the XML document, using these unique XPath expressions to do so.

The next step was to further develop our script to produce tidy data sets, as defined and introduced by @Wickham2014. These contain all parsed text elements, where each statement is separated onto its own row with details about the speaker, and rows are maintained in chronological order. This first involved correcting the variable classes and adding several indicator variables to differentiate where statements came from, such as Chamber versus Federation Chamber or `<sub-debate.1> versus `<sub-debate.2>`. The next key task stemmed from the fact that the raw text data were not separated by each statement when parsed. In other words, any interjections, comments made by The Speaker or Deputy Speaker and continuations within an individual speech were all parsed together as a single string. As such, the name, name ID, electorate and party details were only provided for the person whose turn it was to speak. Further details on the intricacies of the task of splitting these speeches in a way that would be generalizable across sitting days are provided in @sec-interject.

Since we are looking at a wide time span of documents, there are many changes in the way they are formatted. These became apparent as we ran our script on XML files from earlier sitting days. Some changes are as subtle as a differently named child-node, while others are as extensive as a completely different nesting structure. Smaller changes were accounted for as we became aware of them, and embedded into the code in a way that would not cause issues for parsing more current Hansards with subtle differences in formatting. However, as mentioned, more significant changes in the XML structure of Hansard are what necessitated the development of separate scripts as we worked backwards. Further not every sitting day contains every possible XML element. For example, some days did not have `<sub-debate.2>` content, and some days did not have a Federation Chamber proceeding. To improve the generalizability of these scripts, if-else statements were embedded within the code wherever an error might arise due to a missing element. For example, the entire Federation Chamber block of code is wrapped in an if-else statement for each script, so that it only executes if what the code attempts to parse exists in the file.

Once the script ran without error for a few recent years of Hansard, we continued to work backwards until extensive changes in tree structure made our script incompatible with parsing earlier XML files. The earliest sitting day this first script can successfully parse is 14 August 2012. Before developing new scripts to parse earlier Hansard documents, we prioritized cleaning and finalizing what we had been able to parse. As such we continued building our script, fixing any problems we noticed in the resulting datasets such as excess whitespace or spacing issues, and splitting up any additional sections of the parsed text onto separate rows where necessary. Specifically, we added a section of our script to separate out general stage directions. More information on this separation will be provided in @sec-stage. After completing our first script, it was formatted as a function which takes a single file name argument and produces one CSV file containing data on all proceedings from the given sitting day.

@@

As mentioned, we developed a total of four scripts to parse the 1998-2022 time frame of Hansard documents. Two main factors motivated us to create four scripts as opposed to just one, the first being structural variation in XML over time, and the second being improved computational efficiency with separate scripts. While all four scripts use the same general approach to parsing described in @sec-overview and produce the same CSV structure, the first and second scripts use a different method of data processing than the third and fourth scripts.

@@

While the modifications needed to develop the second script were straightforward, this was not the case for our next script. The typical tree structure of Hansard XMLs spanning from 1998 to March 2011 has an important difference from that of XMLs released after March 2011, necessitating many changes to be made in our methodology. In XMLs after March 2011, which our first two scripts successfully parse, the first two child-nodes of `<speech>` are typically `<talk.start>`, and `<talk.text>`. The first child-node contains data on the person whose turn it is to speak, and the second contains the entire contents of that speech --- including all interjections, comments, and continuations. After the `<talk.text>` element closes, there are typically a series of other child-nodes which provide a skeleton structure for how the speech proceedings went in chronological order. For example, if the speech began, was interrupted by one member, and then continued uninterrupted until the end, there would be one `<interjection>` node and one `<continuation>` node following the `<talk.text>` node. These would contain details on the person who made each statement, such as their party and electorate.

@@

This was not isolated to just the beginning of speeches --- details on individuals interjecting or commenting during speeches were also captured this way. To separate statements correctly, we collected all of these patterns using the `<talk.start>` node, and used them to split statements wherever one of these patterns was found. After separating the statements, we were able to remove these patterns from the body of text. We also used this method of extracting and later removing unwanted patterns for other pieces of data which did not belong to the debate proceedings, such as sub-debate titles.

@@

Our approach to parse the `"chamber.xscript"` speeches used in all four scripts meant that all questions without notice content was already parsed in order. For the first two scripts, questions and answers were already separated onto their own rows. For the third and fourth scripts, just as we did with the rest of the speech content, we used those patterns of data preceding the text to separate questions and answers. Finally, since questions in writing exist in their own child node we were able to use the same parsing method for all scripts, which was to extract all question and answer elements from the `"answers.to.questions"` child node.

We then added binary flags to differentiate between questions and answers. To do this in the first and second scripts, we separately re-parsed question and answer content using the XPath expressions `"chamber.xscript//question"` and `"chamber.xscript//answer"`, added the correct question and answer flags accordingly, and then added those flags back to the main dataframe based on exact text matches. For the third and fourth scripts, we made use of the fact that the patterns preceding text transcribed under a question node were stored separately from those transcribed under an answer node. As a result, we could readily use those patterns to flag questions and answers correctly based on which list of patterns it belonged to. Sometimes, questions were incorrectly transcribed under an answer node and vice-versa, in which cases we manually corrected the question and answer flags. For instance, we check for any statements flagged as questions which include the phrase "has provided the following answer to the honourable member's question", in which case we re-code that statement as an answer.

The next step was to merge Question Time contents with all the debate speech. Our method of parsing meant that everything was already in order, so we did not have to perform any additional merging. For questions in writing, merging this content was also straightforward due to the fact that it is always at the end of Hansard. This means that we could bind question in writing rows to the bottom of the main dataframe. This approach was used for all four scripts.

@@

Below is an example of part of a speech we would need to split, extracted from Hansard on 30 November 2021, where Bert van Manen is interrupted by the Speaker who states that the time for members' statements has concluded.

@@

As mentioned in @sec-diff, more recent years of Hansard XMLs contain a series of child-nodes which exist to capture the structure of interruptions in that speech. @fig-xml3 provides an example of this, where the speech was interrupted by a comment from the Deputy Speaker, and then the member continued their speech. Looking at the element names highlighted in blue, these child-nodes do not contain the actual text for the interjection or continuation --- this text is embedded within the speech above it. However, as shown by the content highlighted in pink in @fig-xml3, we were able to extract useful details on the individual interjecting which we could use later. Making use of this structure, we extracted names and information of all individuals that were categorized within the XML as interjections. We stored this as a dataframe called `"interject"`. We decided not to include this data in our final database, as it is embedded in our resulting datasets which have a flag for interjections.

@@

We then created lists using both the `interject` and `main` dataframes to capture all the names of individuals who spoke that day. We added the names of all members in a number of unique formats, due to the frequent variation in how names are transcribed in Hansard. When a member interjects or continues a speech, the usual form of their name is a title followed by their first name or first initial and/or last name. There is also variation in the capitalization of these names.[^2] Another source of variation is in individuals with more than one first name, as sometimes only their initial first name is written, while other times their entire first name is written. Additionally, some surnames have punctuation, and some surnames have specific capitalization such as "McCormack", where even in full capitalization, the first "c" remains lower case. This variation demands careful consideration when writing regular expression patterns. In these lists we also accounted for any general interjection statements that were not attributed to an individual, such as "An opposition member interjecting-".

@@

Having these lists enabled us to extract the names of members and their associated titles as they exist in the text, by searching for exact matches with regular expression patterns. We then used these extracted names to split all the speeches, using regular expressions with lookarounds. A lookaround can be added to a regular expression pattern to enhance the specificity of matches. These were used to ensure that the text was not being split in the wrong places, such as places where members were being named in the statement of another member.

@@

Next, we merged our main dataframe with the lookup table to replace any incomplete names with their full names, and to fill in any gaps with available name ID, electorate, party, gender, and unique ID information. Finally, we were able to add a flag for interjections. Grouping our data by the speech number, we defined an interjection as a statement made by anyone who is not the Speaker, the Deputy Speaker, or the member whose turn it was to speak. @fig-interject provides an example of a Federation Chamber proceeding with interjections. Statements made by the member whose turn it was to speak, or by the Deputy Speaker Maria Vamvakinou, are not flagged as interjections.

@@

Having developed a successful methodology for splitting interjections, we used this to inform our general approach in the third and fourth scripts. However, the difference in data cleaning used in these scripts necessitated some departure from the original methodology. As discussed in @sec-diff, we used string patterns extracted from `<talk.start>` nodes to separate speeches. As evident in @fig-xml3, `<talk.start>` nodes are nested within `<interjection>` nodes, meaning that the patterns of data from interjection statements were separated out in the process. In other words, our approach to data cleaning in the third and fourth scripts separated out interjections in the process. This meant that we did not need to create lists of names and titles for which to search in the text as we did before. However, we used the same list of general interjection statements on which to separate as was used in the first two scripts. We then did an additional check for statements that may have not been separated due to how they were embedded in the XML, and separated those out where needed.[^3]

@@

# Next Steps

While our database makes a substantial contribution to the field, there are a few areas for future development.

One important contribution would be to take advantage of similar records that exist for the Senate over this same period and add them to our database. The main challenge of this would be accounting for different parliamentarians, but the broad structures would be similar.

Earlier House of Representatives records also exist. In some cases it is difficult to get XML records for certain dates, and the earlier decades would require more cleaning.
